quollr: An R Package for Visualizing 2-D Models from Non-linear Dimension Reductions in High Dimensional Space

### 2025-05-19 paper-quollr.Rmd
HTML version is working
Added figure position

---
title: "quollr: An R Package for Visualizing 2-D Models from Non-linear Dimension Reductions in High Dimensional Space"
abstract: >
  Non-linear dimension reduction (NLDR) methods provide a low-dimensional representation of high-dimensional data ($p\text{-}D$) by applying a non-linear transformation. However, the complexity of the transformations and data structures can create wildly different representations depending on the method and (hyper-)parameter choices. It is difficult to determine whether any of these representations are accurate, which one is the best, or whether they have missed important structures. The R package \CRANpkg{quollr} has been developed as a new visual tool to determine which method and which (hyper-)parameter choices provide the most accurate representation of high-dimensional data. The `scurve` data from the package is used to illustrate the algorithm. Single-cell RNA sequencing (scRNA-seq) data from mouse limb muscles are used to demonstrate the usability of the package.
draft: true
author:  
  - name: Jayani P. Gamage
    affiliation: Monash University
    address: Department of Econometrics and Business Statistics, VIC 3800 Australia
    url: https://jayanilakshika.netlify.app/
    orcid: 0000-0002-6265-6481
    email:  \email{jayani.piyadigamage@monash.edu}
  - name: Dianne Cook
    affiliation: Monash University
    address: Department of Econometrics and Business Statistics, VIC 3800 Australia
    url: http://www.dicook.org/
    email: dicook@monash.edu
    orcid: 0000-0002-3813-7155
  - name: Paul Harrison
    affiliation: Monash University
    address: MGBP, BDInstitute, VIC 3800 Australia
    email: paul.harrison@monash.edu
    orcid: 0000-0002-3980-268X
  - name: Michael Lydeamore
    affiliation: Monash University
    address: Department of Econometrics and Business Statistics, VIC 3800 Australia
    email: michael.lydeamore@monash.edu
    orcid: 0000-0001-6515-827X
  - name: Thiyanga S. Talagala
    affiliation: University of Sri Jayewardenepura
    address: Department of Statistics, Gangodawila, Nugegoda 10100 Sri Lanka
    url: https://thiyanga.netlify.app/
    email: ttalagala@sjp.ac.lk
    orcid: 0000-0002-0656-9789
type: package
creative_commons: CC BY
date: "`r Sys.Date()`"
preamble: >
  \usepackage{amsmath}
  \usepackage{array}
output: 
 rjtools::rjournal_article:
    keep_md: true
bibliography: paper-quollr.bib
editor_options: 
  chunk_output_type: inline
---

@@

@@

```{r, eval=knitr::is_latex_output()}
hex_grid_scurve + wrap_plots(
  scurve_proj_umap_model1, scurve_proj_umap_model2,
  scurve_proj_umap_model3, scurve_proj_umap_model4, 
  ncol = 2)
```

@@

```{r, eval=knitr::is_latex_output(), out.width="100%", fig.height=5, fig.width=25}
scurve_umap_plt + hex_grid_poly_scurve +
  hex_centroids_scurve + wireframe_scurve + 
  scurve_proj_umap_model1 +
  plot_layout(ncol = 5)
```

>>>>

---
title: "quollr: An R Package for Visualizing 2-D Models from Non-linear Dimension Reductions in High Dimensional Space"
abstract: >
  Non-linear dimension reduction (NLDR) methods provide a low-dimensional representation of high-dimensional data ($p\text{-}D$) by applying a non-linear transformation. However, the complexity of the transformations and data structures can create wildly different representations depending on the method and (hyper-)parameter choices. It is difficult to determine whether any of these representations are accurate, which one is the best, or whether they have missed important structures. The R package \CRANpkg{quollr} has been developed as a new visual tool to determine which method and which (hyper-)parameter choices provide the most accurate representation of high-dimensional data. The `scurve` data from the package is used to illustrate the algorithm. Single-cell RNA sequencing (scRNA-seq) data from mouse limb muscles are used to demonstrate the usability of the package.
draft: true
author:  
  - name: Jayani P. Gamage
    affiliation: Monash University
    address: Department of Econometrics and Business Statistics, VIC 3800 Australia
    url: https://jayanilakshika.netlify.app/
    orcid: 0000-0002-6265-6481
    email:  \email{jayani.piyadigamage@monash.edu}
  - name: Dianne Cook
    affiliation: Monash University
    address: Department of Econometrics and Business Statistics, VIC 3800 Australia
    url: http://www.dicook.org/
    email: dicook@monash.edu
    orcid: 0000-0002-3813-7155
  - name: Paul Harrison
    affiliation: Monash University
    address: MGBP, BDInstitute, VIC 3800 Australia
    email: paul.harrison@monash.edu
    orcid: 0000-0002-3980-268X
  - name: Michael Lydeamore
    affiliation: Monash University
    address: Department of Econometrics and Business Statistics, VIC 3800 Australia
    email: michael.lydeamore@monash.edu
    orcid: 0000-0001-6515-827X
  - name: Thiyanga S. Talagala
    affiliation: University of Sri Jayewardenepura
    address: Department of Statistics, Gangodawila, Nugegoda 10100 Sri Lanka
    url: https://thiyanga.netlify.app/
    email: ttalagala@sjp.ac.lk
    orcid: 0000-0002-0656-9789
type: package
creative_commons: CC BY
date: "`r Sys.Date()`"
preamble: >
  \usepackage{amsmath}
  \usepackage{array}
  \usepackage{float}
output: 
 rjtools::rjournal_article:
    keep_md: true
bibliography: paper-quollr.bib
editor_options: 
  chunk_output_type: inline
---

@@

```{r, eval=knitr::is_html_output(), fig.cap="algorithm"}

@@

```{r, eval=knitr::is_latex_output(), fig.cap="algorithm", fig.pos='H'}
hex_grid_scurve + wrap_plots(
  scurve_proj_umap_model1, scurve_proj_umap_model2,
  scurve_proj_umap_model3, scurve_proj_umap_model4, 
  ncol = 2)
```

@@

```{r, eval=knitr::is_latex_output(), out.width="100%", fig.height=5, fig.width=25, fig.pos='H'}
scurve_umap_plt + hex_grid_poly_scurve +
  hex_centroids_scurve + wireframe_scurve + 
  scurve_proj_umap_model1 +
  plot_layout(ncol = 5)
```

### 2025-05-19 paper-quollr.tex
Fixed the issue of rendering pdf (removed \maketitle)
Added figure position

\includegraphics[width=1\linewidth]{paper-quollr_files/figure-latex/unnamed-chunk-3-1}

@@

>>>>

\begin{figure}[H]
\includegraphics[width=1\linewidth]{paper-quollr_files/figure-latex/unnamed-chunk-3-1} \caption{algorithm}\label{fig:unnamed-chunk-3}
\end{figure}

@@

\includegraphics[width=1\linewidth]{paper-quollr_files/figure-latex/unnamed-chunk-31-1}
